---
title: "OPSS"
author: "Jason Scheer"
date: "2024-07-11"
output: pdf_document
---


In Baseball the most common advanced metric to use to measure offensive value is OPS (On-Base Plus Slugging) or OPS+ and it seeks to show the value of a hitter at the plate by taking their OBP (On-Base Percentage) and their SLG (Slugging Percentage) and combining them together to combine the 2 most important parts of being a hitter: getting on base and how many bases they get when they get a hit. I think it's a good statistic to use to get general offensive value, but I think there is one key element missing from this offensive value: Stolen Bases. If you look at 2024 Elly De La Cruz and Salvador Perez, both have a similar OPS (as of 7/10/24). But should a single count the same for each of them? How about if I said that Elly has 45 SB and Perez has 0 SB. Shouldn't a single or a double mean more for Elly if he can just get an extra base? Why should a standup double and a single with a stolen base be viewed very differently. We should reward the double more than a single with a stolen base, but OPS was made to determine offensive performance and I think that another statistic should be made to, not punish those who don't steal, but to reward those who do by a certain factor. Let's take a look and see how we would do this.


I have two different formulas to try this new statistic out. It can either increase by the difference of SB and CS and everyone, for the most part, will have their number increase or it can be one that takes run value of getting CS into account. My first formula is:


OPS * ((AB + SB - CS) / AB)


My second formula is:


OPS * (AB + 3(SB) - 4(CS) / AB)


Both I think are good formulas, I think it more depends on what the difference between players will be and if either will favor speedsters too much

```{r}
library(readr)
library(dplyr)
library(stringi)
library(tidyr)

Batting <- read_csv("~/Desktop/lahman_1871-2023_csv/lahman_1871-2023_csv/Batting.csv", show_col_types = FALSE)
batting_2023 = Batting |> 
  filter(yearID == '2023') |> 
  filter(G_batting > 0) |> 
  filter(AB >= 50) |> # Arbitrary number to get rid of players who will skew data
  filter(AB >= (3.1 * G)) # Qualified Batters only
```


```{r}
# Get names of every player so looking at names will be easier later
People = read_csv("~/Desktop/lahman_1871-2023_csv/lahman_1871-2023_csv/People.csv", show_col_types = FALSE)
names = People |> 
  mutate(Name = paste(nameFirst, nameLast)) |> 
  mutate(Name = stri_trans_general(Name, "Latin-ASCII")) |> 
  select(c(playerID, Name))

# To keep this simple for OPS+ calculations, I'm only going to use a players' first team he played on
batting_2023 = batting_2023 |> 
  filter(stint == '1')


# Combine data frames to get Names into Batting stats
batting_2023 = batting_2023 |> 
  left_join(names, by = "playerID") |> 
  select(c("playerID", "Name", "teamID", "G", "AB", "R", "H", "2B", "3B", "HR", "RBI", "SB", "CS", "BB", "SO", "IBB", "HBP", "SH", "SF", "GIDP"))
```



```{r}
# added variables to calculate AVG, SLG, and OPS
batting_2023 = batting_2023 |> 
  mutate(`1B` = (H - `2B` - `3B` - HR)) |> 
  replace_na(list(HBP = 0, SF = 0)) |> 
  mutate(
    OBP = round(((H + BB + HBP) / (AB + BB + HBP + SF)), digits = 3),
    SLG = round(((`1B` + (2 * `2B`) + (3 * `3B`) + (4 * HR)) / AB), digits = 3),
    SLG_3 = round(((SB + (-1.5 * CS) + `1B` + (2 * `2B`) + (3 * `3B`) + (4 * HR)) / AB), digits = 3), 
    OPS = OBP + SLG)
```


Now that we have OPS for each player, let's try adding our two OPSS calculations. I'll call OPSS1 and OPSS2 respectively along with adding OPS+, OPSS1+, and OPSS2+

To calculate OPS, I will be using 0.734 OPS as the league average (via baseball reference) and park factors from Baseball Savant


```{r}
park_factors = data.frame(
  teamID = c("COL", "BOS", "CIN", "KCA", "WAS", "TEX", "PIT", "ATL", "CHN", "PHI", "BAL", "ARI", "MIN", "LAA", "TOR", "HOU", "CHA", "SLN", "LAN", "NYA", "MIA", "SFN", "DET", "CLE", "MIL", "NYN", "TBA", "OAK", "SDN", "SEA"),
  park_factor = c(112, 108, 107, 104, 103, 102, 101, 101, 101, 101, 101, 100, 100, 100, 100, 100, 100, 99, 99, 98, 98, 97, 97, 97, 97, 96, 96, 96, 95, 92)
)

batting_2023 = batting_2023 |> 
  left_join(park_factors, by = "teamID")

batting_2023 = batting_2023 |> 
  mutate(`OPS+` = round( 10000 * (OPS / 0.734) / park_factor, digits = 0 ) ) |> 
  arrange(desc(`OPS+`))

head(batting_2023, n = 10)
```


Now it's time to add each formulas and then the '+' of each formula once I find the league average for each

OPS * ((AB + SB - CS) / AB)

OPS * (AB + 3(SB) - 4(CS) / AB)


```{r}
opss = batting_2023 |> 
  mutate(OPSS1 = round(OPS * ((AB + SB - CS) / AB), digits = 3)) |> 
  mutate(OPSS2 = round(OPS * ((AB + (3 * SB) - (4 *CS)) / AB), digits = 3)) |> 
  mutate(OPSS3 = round(OBP + SLG_3, digits = 3))

avg_opss1 = round(mean(opss$OPSS1), digits = 3)
avg_opss1
avg_opss2 = round(mean(opss$OPSS2), digits = 3)
avg_opss2
avg_opss3 = round(mean(opss$OPSS3), digits = 3)
avg_opss3
```



```{r}
opss = opss |> 
  mutate(`OPSS1+` = round( 10000 * (OPSS1 / avg_opss1) / park_factor, digits = 0 ) ) |>
  mutate(`OPSS2+` = round( 10000 * (OPSS2 / avg_opss2) / park_factor, digits = 0 ) ) |> 
  mutate(`OPSS3+` = round( 10000 * (OPSS3 / avg_opss3) / park_factor, digits = 0 ) ) |> 
  select(c("Name", "teamID", "park_factor", "G", "AB", "SB", "CS", "OBP", "SLG", "OPS", "OPS+", "OPSS1", "OPSS1+", "OPSS2", "OPSS2+", "OPSS3", "OPSS3+"))

head(opss, n = 10)
```

Here our 10 top of highest OPSS+: Ohtani, Judge, Acuna, Schneider, Betts, Alvarez, Seager, Freeman, Olson, and Soto

Lets sort by OPSS1 and OPSS2 to see if there's any change in the top 10


```{r}
opss1 = opss |> 
  arrange(desc(`OPSS1+`))

head(opss1, n = 10)
```

Here are the top 10 for OPSS1+: Ohtani, Acuna, Judge, Schneider, Betts, Freeman, Seager, Alvarez, Olson, Soto



```{r}
opss2 = opss |> 
  arrange(desc(`OPSS2+`))

head(opss2, n = 10)
```

Here are the top 10 for OPSS2+: Acuna, Ohtani, Freeman, Carroll, Judge, Betts, Schneider, Lowe, Soto, and Seager



```{r}
opss3 = opss |> 
  arrange(desc(`OPSS3+`))

head(opss3, n = 10)
```





So if we compare the top 10 of each of the three different measurements we can see that OPS+ and OPSS1+ have the same top 10 just a little rearranged while OPSS2+ has Carroll and Lowe jumping into the top 10 and Alvarez and Olson falling out. So it seems that OPSS1+ doesn't put as much emphasis on stolen bases as compared with OPSS2+ just by looking at the top 10 with OPSS1+ having a different order but the top 10 hitters all staying up in the top 10 while OPSS2+ had Carroll jump not just into the top 10 but into #4 on the list. OPSS2+ also varied a little differently than the other two at the top with having a massive outlier with Acuna at the top. It also seems like more players will be closer to average with adding this in which makes since players who can hit better but don't run well or players who don't hit well but run well will average out a little more. Now that we see how it affects the bottom 10 and the middle("average") players. 


```{r}
opss = opss |> 
  arrange(`OPS+`)

head(opss, n = 10)
```



```{r}
opss1 = opss1 |> 
  arrange(`OPSS1+`)

head(opss1, n = 10)
```



```{r}
opss2 = opss2 |> 
  arrange(`OPSS2+`)

head(opss2, n = 10)
```


```{r}
opss3 = opss3 |> 
  arrange(`OPSS3+`)

head(opss3, n = 10)
```


The bottom 10 look relatively the same. We see about the same amount of variance in OPSS1+ and OPSS2+. I'm really interested to see how the average player is affected by each variable.


```{r}
median(opss$`OPS+`)
median(opss$`OPSS1+`)
median(opss$`OPSS2+`)
median(opss$`OPSS3+`)
```


Here we see that the median is brought down by both of them and that OPSS2+ brings it down below 100, the "average" player. This makes sense since there will be more bad hitters being brought up with speed or good hitters being brought down with bad baserunning.



```{r}
ops_mean = mean(opss$`OPS+`, na.rm = TRUE)
ops_mean
opss1_mean = mean(opss$`OPSS1+`, na.rm = TRUE)
opss1_mean
opss2_mean = mean(opss$`OPSS2+`, na.rm = TRUE)
opss2_mean
opss3_mean = mean(opss$`OPSS3+`, na.rm = TRUE)
opss3_mean
ops_sd = sd(opss$`OPS+`, na.rm = TRUE)
ops_sd
opss1_sd = sd(opss$`OPSS1+`, na.rm = TRUE)
opss1_sd
opss2_sd = sd(opss$`OPSS2+`, na.rm = TRUE)
opss2_sd
opss3_sd = sd(opss$`OPSS3+`, na.rm = TRUE)
opss3_sd
```


```{r}
# 66% and 95% of players deviation for OPS+
c(ops_mean - ops_sd, ops_mean + ops_sd)
c(ops_mean - (2 * ops_sd), ops_mean + (2 * ops_sd))

cat("
")
# 66% and 95% of players deviation for OPS+
c(opss1_mean - opss1_sd, opss1_mean + opss1_sd)
c(opss1_mean - (2 * opss1_sd), opss1_mean + (2 * opss1_sd))

cat("
")
# 66% and 95% of players deviation for OPS+
c(opss2_mean - opss2_sd, opss2_mean + opss2_sd)
c(opss2_mean - (2 * opss2_sd), opss2_mean + (2 * opss2_sd))

cat("
")
# 66% and 95% of players deviation for OPS+
c(opss3_mean - opss3_sd, opss3_mean + opss3_sd)
c(opss3_mean - (2 * opss3_sd), opss3_mean + (2 * opss3_sd))
```

Here we can see the comparisons of the means and standard deviations to see where 66% and 95% of batters lie between and see how they differ. Each statistic increased in range on the low end while OPSS1+ is shorter than the other two on the high end meaning that there isn't as many high end players in OPSS1+ as OPS+ and OPSS2+




Lastly I want to look into how OPSS1+ and OPSS2+ differ from OPS+ from each player and see if there's anything interesting to find there


```{r}
opss = opss |> 
  mutate(`OPSS1+_diff` = abs(`OPS+` - `OPSS1+`)) |> 
  mutate(`OPSS2+_diff` = abs(`OPS+` - `OPSS2+`)) |> 
  mutate(`OPSS3+_diff` = abs(`OPS+` - `OPSS3+`))

opss1 = opss |> 
  arrange(desc(`OPSS1+_diff`))

head(opss1, n = 15)
```


```{r}
opss2 = opss |> 
  arrange(desc(`OPSS2+_diff`))

head(opss2, n = 15)
```


```{r}
opss3 = opss |> 
  arrange(desc(`OPSS3+_diff`))

head(opss3, n = 15)
```





Now this is, to me, the most interesting thing to look at. Here it shows the absolute value difference in OPS+ to OPSS1+ and OPSS2+ and this is where we see the biggest difference between these 3 variables. In OPSS1+ you see some big differences in 7 and 6, but not too much big differences, but OPSS2+ is where we see the biggest of differences. 4 players with over 20 differential which means this stat made those 20% better/worse than they were with OPS+. This is where the decision between models for OPSS is made. Depending on how much you want stolen bases to affect their offensive metric, you can have it impact it a little bit but not differ a ton or you have it mean a decent amount with a few huge outliers, but most still having a bigger affects than OPSS1+ by a mile. 



Summary: 

Even though we see a lot of similarities and each model having its own pros and cons, I think OPSS1+ is a better model purely down to the power stolen bases adds to offensive production. I think it shoud matter, but I don't think it should be able to sway an offensive statistic has much as it does. 
